---
title: "Data Modelling 2011-2012 Data Set"
author: "Niamh Harford 20251644"
date: "5/7/2021"
output: html_document
---



```{r}
# Load in all the data
load("~/Documents/Masters/Dissertation/Project_Git/Data/data1_11a.RData") 

head(data_11a)
nrow(data_11a)
# Minus Don't Know and Nas 2011-2012
```


```{r}
# Load in necessary packages
suppressMessages(library(tidyverse))
suppressMessages(library(sjPlot))
suppressMessages(library(sjlabelled))
suppressMessages(library(sjmisc))
suppressMessages(library(MASS))
```



##### Data Minus Don't Know and NA Values 2011-2012

```{r, results='hide'}
#Logistic regression
data_11a$Hypertension <-ifelse(data_11a$Hypertension == "Yes", 1 ,0)

data_11a$Hypertension %>% table

log_reg1 <- glm(Hypertension ~ ., data = data_11a, family = binomial())
summary(log_reg1)
```

```{r}
coefficients(log_reg1) %>% exp() %>% round(digits = 5)
```

#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to decrease by roughly 9%. 

+ For a one unit increase in age you would expect your odds of having hypertension to increase by roughly 6%.

+ When you move from American Mexican race to Other Hispanic, the odds of having hypertension increases by roughly 26%.

+ When you move from American Mexican race to Non-Hispanic White, the odds of having hypertension increases by roughly 33%.

+ When you move from American Mexican race to Non-Hispanic Black, the odds of having hypertension is expected to be roughly 2.21 times higher.

+ When you move from American Mexican race to Other Race including multi-racial, the odds of having hypertension increases by roughly 31%.

+ When you move from not smoking to smoking, the odds of having hypertension increases by roughly 24%.

+ When you move from having not diabetes to having diabetes, the odds of having hypertension increases by roughly 84%.

+ When you move from having not diabetes to borderline having diabetes, the odds of having hypertension is expected to be roughly 2.45 times higher.

+ When you move from taking not insulin to taking insulin, the odds of having hypertension increases by roughly 18%.

+ When you move from having chronic kidney disease to not having chronic kidney disease, the odds of having hypertension is expected to be roughly 2.86 times higher.

+ For a one unit increase in BMI you would expect your odds of having hypertension to increase by roughly 8%. 

+ When you move from not being physically active to being physically active, the odds of having hypertension decreases by roughly 10%.

+ When you move from drinking at least 12 alcoholic drinks in a year to not, you would expect your odds to have no change.

+ When you move from never using salt to rarely, the odds of having hypertension decreases by roughly 13%.

+ When you move from never using salt to occasionally, the odds of having hypertension increases by roughly 5%.

+ When you move from never using rarely to very often, you would expect your odds to have no change.

+ For a one unit increase in potassium you would expect your odds of having hypertension to have no change.


**Variables that increases the odds of Hypertension **

+ 1 Unit Increase in Age
+ Gender
+ Smoking 
+ Having Diabetes
+ Having Borderline Diabetes
+ Taking Insulin
+ Having Chronic Kidney Disease
+ 1 Unit Increase in BMI
+ Drinking At Least 12 Alcoholic Drink in a Year
+ Occassionally Using Salt in Food Prep



**Plotting estimates of logistic regression models**

```{r}
#2011-2012 Reduced Data
plot_model(log_reg1, vline.color = "black", sort.est = TRUE,colors = "Accent")
```





**Residuals**


```{r}
plot(log_reg1)
# Potential outliers : 857,1537,2637
```

```{r}
# 2011-2012 All Data
suppressMessages(library(arm))
binnedplot(fitted(log_reg1), 
           residuals(log_reg1, type = "response"), 
           nclass = NULL, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = "blue", 
           col.int = "black")

# The grey lines represent  ±  2 SE bands, which we would expect to contain about 95% of the observations.
# https://bookdown.org/jefftemplewebb/IS-6489/logistic-regression.html 
# The binned residuals plot instead, after dividing the data into categories (bins) based on their fitted values, the average residual versus the average fitted value for each bin
```

```{r}
suppressMessages(library(car))
residualPlots(log_reg1)
# a lack-of-fit test is provided only for the numeric predictor
# https://www.sagepub.com/sites/default/files/upm-binaries/38503_Chapter6.pdf
# Age and BMI plots may indicate lack of fit


influenceIndexPlot(log_reg1, vars=c("Cook", "hat"), id.n=3)
# Observations 857, 1537 and 2637 have the largest Cook's distance, although even these are quite small. I shall remove the 3 cases as a check

compareCoefs(log_reg1, update(log_reg1, subset=-c(857,1537,2637)))
# Removing these 3 observations does not alter the results much at all
```


```{r}
prob <- predict(log_reg1, type="response")
pred <- factor(ifelse(prob < .5, "No", "Yes"))
table(data_11a$Hypertension, pred)

# Error rate
(651+440)/(2296+651+440+925)*100


# MSE
mean(log_reg1$residuals^2)
```




**Smoothing Spline on BMI and Age**

```{r, eval=FALSE,echo=FALSE}
# head(data_11a)
# # Hypertension vs BMI
# suppressMessages(library(splines))
# 
# f1 <- smooth.spline(data_11a$BMI,data_11a$Hypertension, cv = T)
# # cross-validation with non-unique 'x' values seems doubtful
# 
# ggplot(data=data_11a, aes(x=BMI, y=Hypertension))+ geom_point() +
# geom_line(aes(y=fitted(f1)), color="green")
# 
# f2 <- glm(Hypertension ~ bs(BMI,df=5), data=data_11a, family = binomial)
# attr(bs(data_11a$BMI, df=5), "knots")
# 
# ggplot(data=data_11a, aes(x=BMI, y=Hypertension)) + geom_point() + 
#   geom_line(aes(y=fitted(f2)), color= "blue") + geom_vline(xintercept = attr(bs(data_11a$BMI, df=5), "knots"), colour = "magenta")

# ggplot(data_11a) +
#       geom_bar(aes(x = Diabetes, fill = Hypertension),position = "dodge") + 
#       facet_wrap(~Race) + xlab("Smoking") + labs(fill= "Hypertension")
```

```{r}
suppressMessages(library(gam))
gam_logistic = gam(Hypertension ~ Gender + bs(Age, df = 5) + Race + Smoking + Diabetes + 
                     Insulin + KidneyF + bs(BMI,df=5) + PhysActivity + Alcohol + Salt +
                     Potassium, data=data_11a, family = binomial())

plot(gam_logistic, se = T, col = "green")
# 1st plot; males (0.5) have a higher risk of hypertension than females (-0.5)
# 2nd plot; age has somewhat of a linear term
# 8th plot; bmi doesn't look too different to linear
# The spline terms may not be necessary
# 12th plot; there seems to be no effect of potassium 
```

```{r}
# Comparing Gam and Logistic regression
anova(log_reg1,gam_logistic)

# The GAM model seems to fit better

AIC(log_reg1)
AIC(gam_logistic)
```



```{r, eval=FALSE,echo=FALSE}
#Decision Trees - Doesn't show all variables

suppressMessages(library(tree))
tree <- tree(Hypertension ~ BMI + Age, data=data_11a)
summary(tree)
plot(tree)
text(tree, cex=.5, pretty=0)

tree1 <- tree(Hypertension ~ ., data=data_11a)
summary(tree1)
plot(tree1)
text(tree1, cex=.5, pretty=0)

suppressMessages(library(rpart))
f1 <- rpart(Hypertension ~ ., data=data_11a)
plot(f1)
text(f1, cex=.5, pretty = 0)
```




```{r, results='hide'}
# Logistic Regression - Interaction with Gender
load("~/Documents/Masters/Dissertation/Project_Git/Data/data1_11a.RData")

data_11a$Hypertension <-ifelse(data_11a$Hypertension == "Yes", 1 ,0)
data_11a$Hypertension %>% table
log_reg_gender <- glm(Hypertension ~ .*Gender, data = data_11a, family = binomial())
summary(log_reg_gender)
```

```{r}
coefficients(log_reg_gender) %>% exp() %>% round(digits = 4)
```

#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to decrease by roughly 73%. 

+ For a one unit increase in age if you are male, you would expect your odds of having hypertension to increase by roughly 5%.

+ When you move from American Mexican race to Other Hispanic if you are male, the odds of having hypertension increases by roughly 51%.

+ When you move from American Mexican race to Non-Hispanic White if you are male, the odds of having hypertension increases by roughly 66%.

+ When you move from American Mexican race to Non-Hispanic Black if you are male, the odds of having hypertension is expected to be roughly 2.64 times higher.

+ When you move from American Mexican race to Other Race including multi-racial, the odds of having hypertension is expected to increase by roughly 48%.

+ When you move from not smoking to smoking if you are male, the odds of having hypertension increases by roughly 22%.

+ When you move from not having diabetes to having diabetes if you are male, the odds of having hypertension increases by roughly 79%.

+ When you move from not having diabetes to borderline having diabetes if you are male, the odds of having hypertension is expected to be roughly 2 times higher.

+ When you move from not taking insulin to taking insulin if you are male, the odds of having hypertension decreases by roughly 27%.

+ When you move from not having chronic kidney disease to having chronic kidney disease if you are male, the odds of having hypertension is expected to be roughly 2 times higher.

+ For a one unit increase in BMI if you are male, you would expect your odds of having hypertension to increase by roughly 7%. 

+ When you move from being physically active to not being physically active if you are male, the odds of having hypertension decreases by roughly 9%.

+ When you move from not drinking at least 12 alcoholic drinks in a year to drinking at least 12 alcoholic drinks in a year if you are male, the odds of having hypertension decreases by roughly 2%.

+ When you move from never using salt to rarely if you are male, the odds of having hypertension decreases by roughly 8%.

+ When you move from never using salt to occasionally if you are male, the odds of having hypertension increases by roughly 3%.

+ When you move from never using rarely to very often if you are male, the odds of having hypertension increases by roughly 8%.

+ For a one unit increase in potassium if you are male, you would expect no change in your odds.

+ For a one unit increase in age if you are female, the odds of having hypertension increases by roughly 2%.

+ When you move from American Mexican race to Other Hispanic if you are female, the odds of having hypertension decreases by roughly 34%.

+ When you move from American Mexican race to Non-Hispanic White if you are female, the odds of having hypertension decreases by roughly 41%.

+ When you move from American Mexican race to Non-Hispanic Black if you are female, the odds of having hypertension decreases by roughly 34%.

+ When you move from American Mexican race to Other Race including multi-racial if you are female, the odds of having hypertension decreases by roughly 24%.

+ When you move from not smoking to smoking if you are female, the odds of having hypertension increases by roughly 8%.

+ When you move from not having diabetes to having diabetes if you are female, the odds of having hypertension increases by roughly 9%.

+ When you move from not having diabetes to borderline having diabetes if you are female, the odds of having hypertension decreases by roughly 44%.

+ When you move from not taking insulin to taking insulin if you are female, the odds of having hypertension decreases by roughly 17%.

+ When you move from not having chronic kidney disease to having chronic kidney disease if you are female, the odds of having hypertension increases by roughly 18%.

+ For a one unit increase in BMI if you are female, you would expect your odds of having hypertension to increase by roughly 2%. 

+ When you move from being physically active to not being physically active if you are female, you would expect the odds to have no change.

+ When you move from not drinking at least 12 alcoholic drinks in a year to drinking at least 12 alcoholic drinks in a year if you are female, the odds of having hypertension increases by roughly 11%.

+ When you move from never using salt to rarely if you are female, the odds of having hypertension decreases by roughly 12%.

+ When you move from never using salt to occasionally if you are female, the odds of having hypertension increases by roughly 4%.

+ When you move from never using salt to very often if you are female, the odds of having hypertension decreases by roughly 16%.

+ For a one unit increase in potassium if you are female, you would expect your odds to have no change.



**Main Findings for this Logistic Regression**

+ Odds of hypertension expected to decrease if you are female (65%)
+ Odds of hypertension move from having diabetes to borderline having diabetes if you are male expected to increase (12%)
+ Odds of hypertension if you are female and any race are expected to decrease, whereas odds of hypertension if you are male and any race are expected to increase
+ Odds of hypertension expected to increase for females when they are not taking insulin, whereas odds of hypertension if you are male and not taking insulin are expected to decrease


```{r}
#Colour-Blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette

load("~/Documents/Masters/Dissertation/Project_Git/Data/data1_11a.RData") 

data_11a %>% ggplot(aes(x = Hypertension, fill = Gender)) +  geom_bar(position = "dodge") + facet_wrap(~Race) + scale_fill_manual(values=cbPalette[c(3,4)])
```

```{r}
#2011-2012 Gender Interaction


plot_model(log_reg_gender, vline.color = "black", sort.est = TRUE, colors = cbPalette[c(3,4)])

par(mfrow=c(1,2))

plot_model(log_reg_gender, sort.est = TRUE, vline.color = "black", colors = cbPalette[c(3,4)], terms = c("Age","Race [Other Hispanic]","Race [Non-Hispanic White]", "Race [Non-Hispanic Black]", "Race [Other Race - Including Multi-Rac]", "Smoking [Yes]", "Diabetes [Yes]", "Diabetes [Borderline]", "Insulin [Yes]", "KidneyF [Yes]", "BMI", "PhysActivity [Yes]", "Alcohol [Yes]", "Salt [Rarely]", "Salt [Occasionally]", "Salt [Very Often]", "Potassium"), title = "Male Odds of Hypertension")

plot_model(log_reg_gender, sort.est = TRUE, vline.color = "black", colors = cbPalette[c(3,4)], rm.terms = c("Age","Race [Other Hispanic]","Race [Non-Hispanic White]", "Race [Non-Hispanic Black]", "Race [Other Race - Including Multi-Rac]", "Smoking [Yes]", "Diabetes [Yes]", "Diabetes [Borderline]", "Insulin [Yes]", "KidneyF [Yes]", "BMI", "PhysActivity [Yes]", "Alcohol [Yes]", "Salt [Rarely]", "Salt [Occasionally]", "Salt [Very Often]", "Potassium"), title = "Female Odds of Hypertension")

#plot_model(log_reg_gender, show.values = TRUE, value.offset = .3, sort.est = T)
```


```{r}
AIC(log_reg1)
AIC(log_reg_gender)
```



```{r}
load("~/Documents/Masters/Dissertation/Project_Git/Data/data1_11a.RData")

suppressMessages(library(randomForest))
suppressMessages(library(e1071))
suppressMessages(library(DALEX))
suppressMessages(library(rms))
suppressMessages(library(nnet))
suppressMessages(library(devtools))
suppressMessages(source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r'))
 
set.seed(1)

```

```{r}
# Random Forest

model_rf <- randomForest(Hypertension ~ .,data_11a, importance = T)
model_rf # out of bag error rate : ~ 26%
varImpPlot(model_rf) 
# age, bmi and diabetes seem to be important variables in terms of impacting on the response, the same outcome was shown in the logistic plot
```


```{r}

# Neural Networks


model_nn <- nnet(Hypertension ~ ., data=data_11a, size=6)
model_nn
head(data_11a$Hypertension)
pred <- predict(model_nn, data_11a, type="class")
head(pred)
tab <- table(data_11a$Hypertension, pred)
tab

#error rate
(tab[1,1]+tab[2,2])/(sum(tab))*100
mean( data_11a$Hypertension != pred)


suppressMessages(require(scales))
suppressMessages(require(reshape))
plot.nnet(model_nn)
```


```{r}
# Logistic Regression
data_11a$Hypertension <-ifelse(data_11a$Hypertension == "Yes", 1 ,0)

data_11a$Hypertension %>% table
model_log1 <- glm(Hypertension ~ ., data = data_11a, family = binomial())
model_log2  <-glm(Hypertension ~ .*Gender, data = data_11a, family = binomial())

```


```{r}
#Comparing the fits

ex_rf <- explain(model_rf, data = data_11a, y = data_11a$Hypertension)
ex_nn <- explain(model_nn, data = data_11a, y = data_11a$Hypertension)
ex_log1<- explain(model_log1, data = data_11a, y = data_11a$Hypertension)
ex_log2 <- explain(model_log2, data = data_11a, y = data_11a$Hypertension)


plot(model_profile(ex_rf, "Gender"),
     model_profile(ex_nn, "Gender"),
     model_profile(ex_log1, "Gender"),
     model_profile(ex_log2, "Gender")) +
  ggtitle("Responses for Gender") 


plot(model_profile(ex_rf, "Age"),
     model_profile(ex_nn, "Age"),
     model_profile(ex_log1, "Age"),
     model_profile(ex_log2, "Age")) +
  ggtitle("Responses for Age")


plot(model_profile(ex_rf, "Race"),
     model_profile(ex_nn, "Race"),
     model_profile(ex_log1, "Race"),
     model_profile(ex_log2, "Race")) +
  ggtitle("Responses for Race")


plot(model_profile(ex_rf, "Smoking"),
     model_profile(ex_nn, "Smoking"),
     model_profile(ex_log1, "Smoking"),
     model_profile(ex_log2, "Smoking")) +
  ggtitle("Responses for Smoking")


plot(model_profile(ex_rf, "Diabetes"),
     model_profile(ex_nn, "Diabetes"),
     model_profile(ex_log1, "Diabetes"),
     model_profile(ex_log2, "Diabetes")) +
  ggtitle("Responses for Diabetes")


plot(model_profile(ex_rf, "Insulin"),
     model_profile(ex_nn, "Insulin"),
     model_profile(ex_log1, "Insulin"),
     model_profile(ex_log2, "Insulin")) +
  ggtitle("Responses for Insulin") 


plot(model_profile(ex_rf, "KidneyF"),
     model_profile(ex_nn, "KidneyF"),
     model_profile(ex_log1, "KidneyF"),
     model_profile(ex_log2, "KidneyF")) +
  ggtitle("Responses for KidneyF")


plot(model_profile(ex_rf, "BMI"),
     model_profile(ex_nn, "BMI"),
     model_profile(ex_log1, "BMI"),
     model_profile(ex_log2, "BMI")) +
  ggtitle("Responses for BMI")


plot(model_profile(ex_rf, "PhysActivity"),
     model_profile(ex_nn, "PhysActivity"),
     model_profile(ex_log1, "PhysActivity"),
     model_profile(ex_log2, "PhysActivity")) +
  ggtitle("Responses for PhysActivity")


plot(model_profile(ex_rf, "Alcohol"),
     model_profile(ex_nn, "Alcohol"),
     model_profile(ex_log1, "Alcohol"),
     model_profile(ex_log2, "Alcohol")) +
  ggtitle("Responses for Alcohol")

plot(model_profile(ex_rf, "Salt"),
     model_profile(ex_nn, "Salt"),
     model_profile(ex_log1, "Salt"),
     model_profile(ex_log2, "Salt")) +
  ggtitle("Responses for Salt")


plot(model_profile(ex_rf, "Potassium"),
     model_profile(ex_nn, "Potassium"),
     model_profile(ex_log1, "Potassium"),
     model_profile(ex_log2, "Potassium")) +
  ggtitle("Responses for Potassium")

```

```{r}
ex_rf <- explain(model_rf, data = data_11a, y = data_11a$Hypertension)
ex_nn <- explain(model_nn, data = data_11a, y = data_11a$Hypertension)
ex_log1<- explain(model_log1, data = data_11a, y = data_11a$Hypertension)
ex_log2 <- explain(model_log2, data = data_11a, y = data_11a$Hypertension)



sv_rf  <- model_profile(ex_rf, variable =  "Gender", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Gender", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Gender", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Gender", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "Age", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Age", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Age", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Age", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "Race", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Race", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Race", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Race", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "Smoking", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Smoking", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Smoking", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Smoking", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)


sv_rf  <- model_profile(ex_rf, variable =  "Diabetes", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Diabetes", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Diabetes", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Diabetes", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "Insulin", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Insulin", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Insulin", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Insulin", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "KidneyF", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "KidneyF", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "KidneyF", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "KidneyF", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)


sv_rf  <- model_profile(ex_rf, variable =  "BMI", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "BMI", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "BMI", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "BMI", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)


sv_rf  <- model_profile(ex_rf, variable =  "PhysActivity", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "PhysActivity", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "PhysActivity", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "PhysActivity", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)


sv_rf  <- model_profile(ex_rf, variable =  "Alcohol", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Alcohol", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Alcohol", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Alcohol", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)


sv_rf  <- model_profile(ex_rf, variable =  "Salt", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Salt", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Salt", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Salt", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)

sv_rf  <- model_profile(ex_rf, variable =  "Potassium", type = "partial")
sv_log1  <- model_profile(ex_log1, variable =  "Potassium", type = "partial")
sv_log2 <- model_profile(ex_log2, variable =  "Potassium", type = "partial")
sv_nn  <- model_profile(ex_nn, variable =  "Potassium", type = "partial")


plot(sv_rf, sv_log1, sv_log2, sv_nn)
```


```{r}
#Stepwise selection
load("~/Documents/Masters/Dissertation/Project_Git/Data/data1_11a.RData") 
data_11a$Hypertension <-ifelse(data_11a$Hypertension == "Yes", 1 ,0)
data_11a$Hypertension %>% table

model <- glm(Hypertension ~., data = data_11a, family = binomial)
step.model <- model %>% stepAIC(trace = FALSE)
step.model
step.model$coefficients %>% exp
```

^ age,race,smoking,diabetes,kidney,bmi and physical activity


```{r}
library(leaps)

all=regsubsets(Hypertension~., data=data_11a, nbest=1, nvmax=10)
info=summary(all)
cbind(info$which, round(cbind(rsq=info$rsq, adjr2=info$adjr2,
                         cp=info$cp, bic=info$bic, rss=info$rss), 5))
```

