---
title: "Modeling the Data"
author: 'Niamh Harford 20251644'
date: "4/27/2021"
output: html_document
---

```{r}
# Load in all necessary data
load("~/Documents/Masters/Dissertation/Project_Git/Data/all_dat.RData") # All Data 2015-2016
load("~/Documents/Masters/Dissertation/Project_Git/Data/data_09_10.RData") # All Data 2009-2010
load("~/Documents/Masters/Dissertation/Project_Git/Data/hyp1.RData") # Minus Don't Know and Nas 2015-2016
load("~/Documents/Masters/Dissertation/Project_Git/Data/data2_09_10.RData") # Minus Don't Know and Nas 2009-2010
```

```{r}
# Load in necessary packages
suppressMessages(library(tidyverse))
```

#### What Have I Left Out of the Data Sets:

I still have both full data sets for both years (2015-2016 and 2009-2010) but for the 2015-2016 data set there are 1453 missing values and for the 2009-2010 data set there are 1371 missing values. When I remove the 'Don't Know' responses, there are 299 missing values in the 2015-2016 data set and there are 530 missing values in the 2009-2010 data set.

So if I use the reduced data sets (removing the don't know responses), for 2015-2016 data I will have 4628 observations and for 2009-2010 data I will have 5132 observations.

+ Removing Don't Know responses
+ Removing NA values
+ Removing Refused responses (only 2) from smoking variable (2015-2016 Reduced Data)


#### Logistic Regression

##### All Data 2015-2016

```{r, results='hide'}
all_dat <- all_dat %>% filter(Hypertension != "Don't know") %>% droplevels()
all_dat$Hypertension <-ifelse(all_dat$Hypertension == "Yes", 1 ,0)
all_dat$Hypertension %>% table
log_reg_all_dat <- glm(Hypertension ~ ., data = all_dat, family = binomial())
summary(log_reg_all_dat)
```

```{r}
coefficients(log_reg_all_dat) %>% exp() %>% round(digits = 6)
```

#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to increase by roughly 4%. 

+ For a one unit increase in age you would expect your odds of having hypertension to increase by roughly 6%. 

+ When you move from American Mexican race to Other Hispanic, the odds of having hypertension increases by roughly 27%.

+ When you move from American Mexican race to Non-Hispanic White, the odds of having hypertension increases by roughly 19%.

+ When you move from American Mexican race to Non-Hispanic Black, the odds of having hypertension is roughly 2.08 times higher.

+ When you move from American Mexican race to Other Race including multi-race, the odds of having hypertension increases by roughly 67%.

+ When you move from smoking to not smoking, the odds of having hypertension decreases by roughly 16%.

+ When you move from smoking to refused smoking, the odds of having hypertension decreases by roughly 0.0001%.

+ When you move from smoking to don't know smoking, the odds of having hypertension increases by roughly 34%.

+ When you move from having diabetes to not having diabetes, the odds of having hypertension decreases by roughly 57%.

+ When you move from having diabetes to borderline having diabetes, the odds of having hypertension decreases by roughly 14%.

+ When you move from taking insulin to not taking insulin, the odds of having hypertension increases by roughly 4%.

+ When you move from taking insulin to don't know, the odds of having hypertension decreases by roughly 61%.

+ When you move from having chronic kidney disease to not having chronic kidney disease, the odds of having hypertension decreases by roughly 66%.

+ When you move from having chronic kidney disease to don't know if you have chronic kidney disease, the odds of having hypertension decreases by roughly 51%.

+ For a one unit increase in BMI you would expect your odds of having hypertension to increase by roughly 7%. 

+ For a one unit increase in physical activity you would expect your odds of having hypertension to increase roughly by 2%.

+ For a one unit increase in alcohol you would expect your odds of having hypertension to decreases roughly by 12%.

+ When you move from never using salt to rarely, the odds of having hypertension decreases by roughly 19%.

+ When you move from never using salt to occasionally, the odds of having hypertension decreases by roughly 27%.

+ When you move from never using salt to very often, the odds of having hypertension decreases by roughly 26%.

+ When you move from never using salt rarely to don't know, the odds of having hypertension decreases by roughly 58%.

+ For a one unit increase in potassium you would expect your odds of having hypertension to be equally likely.


**Variables that decreases the odds of Hypertension **

+ Not Smoking 
+ Not Having Diabetes
+ Borderline Having Diabetes
+ Not Taking Insulin
+ Don't Know Insulin Intake
+ Not Having Chronic Kidney Disease
+ Don't Know Chronic Kidney Disease
+ Alcohol
+ Use of Salt in Food Prep


##### Data Minus Don't Know and NA Values 2015-2016

```{r, results='hide'}
#Logistic regression
hyp1$Hypertension %>% table
log_reg <- glm(Hypertension ~ ., data = hyp1, family = binomial())
summary(log_reg)
```


```{r}
coefficients(log_reg) %>% exp() %>% round(digits = 6)
```

#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to increase by roughly 4%. 

+ For a one unit increase in age you would expect your odds of having hypertension to increase by roughly 6%. 

+ When you move from American Mexican race to Other Hispanic, the odds of having hypertension increases by roughly 26%.

+ When you move from American Mexican race to Non-Hispanic White, the odds of having hypertension increases by roughly 18%.

+ When you move from American Mexican race to Non-Hispanic Black, the odds of having hypertension is roughly 2.07 times higher.

+ When you move from American Mexican race to Other Race including multi-race, the odds of having hypertension increases by roughly 62%.

+ When you move from smoking to not smoking, the odds of having hypertension decreases by roughly 16%.

+ When you move from having diabetes to not having diabetes, the odds of having hypertension decreases by roughly 57%.

+ When you move from having diabetes to borderline having diabetes, the odds of having hypertension decreases by roughly 19%.

+ When you move from taking insulin to not taking insulin, the odds of having hypertension increases by roughly 7%.

+ When you move from having chronic kidney disease to not having chronic kidney disease, the odds of having hypertension decreases by roughly 67%.

+ For a one unit increase in BMI you would expect your odds of having hypertension to increase by roughly 7%. 

+ For a one unit increase in physical activity you would expect your odds of having hypertension to increase by 4%.

+ For a one unit increase in alcohol you would expect your odds of having hypertension to decreases roughly by 13%.

+ When you move from never using salt to rarely, the odds of having hypertension decreases by roughly 19%.

+ When you move from never using salt to occasionally, the odds of having hypertension decreases by roughly 28%.

+ When you move from never using salt to very often, the odds of having hypertension decreases by roughly 26%.

+ For a one unit increase in potassium you would expect your odds of having hypertension to be equally likely.


**Variables that decreases the odds of Hypertension **

+ Not Smoking 
+ Not Having Diabetes
+ Borderline Having Diabetes
+ Not Having Chronic Kidney Disease
+ Alcohol
+ Use of Salt in Food Prep


##### All Data 2009-2010

```{r, results='hide'}
#Logistic regression

data$Hypertension <-ifelse(data$Hypertension == "Yes", 1 ,0)
data$Hypertension %>% table
log_reg_all <- glm(Hypertension ~ ., data = data, family = binomial())
summary(log_reg_all)
```


```{r}
coefficients(log_reg_all) %>% exp() %>% round(digits = 5)
```


#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to increase by roughly 5%. 

+ For a one unit increase in age you would expect your odds of having hypertension to increase by roughly 7%. 

+ When you move from American Mexican race to Other Hispanic, the odds of having hypertension increases by roughly 23%.

+ When you move from American Mexican race to Non-Hispanic White, the odds of having hypertension increases by roughly 13%.

+ When you move from American Mexican race to Non-Hispanic Black, the odds of having hypertension is roughly 2.11 times higher.

+ When you move from American Mexican race to Other Race including multi-race, the odds of having hypertension increases by roughly 77%.

+ When you move from smoking to not smoking, the odds of having hypertension decreases by roughly 12%.

+ When you move from having diabetes to not having diabetes, the odds of having hypertension decreases by roughly 69%.

+ When you move from having diabetes to borderline having diabetes, the odds of having hypertension decreases by roughly 37%.

+ When you move from having diabetes to don't know, the odds of having hypertension decreases by roughly 46%.

+ When you move from taking insulin to not taking insulin, the odds of having hypertension increases by roughly 40%.

+ When you move from having chronic kidney disease to not having chronic kidney disease, the odds of having hypertension decreases by roughly 77%.

+ When you move from having chronic kidney disease to don't know, the odds of having hypertension decreases by roughly 66%.

+ For a one unit increase in BMI you would expect your odds of having hypertension to increase by roughly 8%. 

+ For a one unit increase in physical activity you would expect your odds of having hypertension to increase by roughly 10%.

+ For a one unit increase in alcohol you would expect your odds of having hypertension increases roughly by 3%.

+ When you move from never using salt to rarely, the odds of having hypertension decreases by roughly 24%.

+ When you move from never using salt to occasionally, the odds of having hypertension decreases by roughly 12%.

+ When you move from never using salt to very often, the odds of having hypertension decreases by roughly 28%.

+ When you move from never using rarely to don't know, the odds of having hypertension increases by roughly 56%.

+ For a one unit increase in potassium you would expect your odds of having hypertension to be equally likely.


**Variables that decreases the odds of Hypertension **


+ Not Smoking 
+ Not Having Diabetes
+ Borderline Having Diabetes
+ Not Having Chronic Kidney Disease
+ Use of Salt in Food Prep



##### Data Minus Don't Know and NA Values 2009-2010

```{r, results='hide'}
#Logistic regression

data2$Hypertension <-ifelse(data2$Hypertension == "Yes", 1 ,0)
data2$Hypertension %>% table

log_reg1 <- glm(Hypertension ~ ., data = data2, family = binomial())
summary(log_reg1)

```

```{r}
coefficients(log_reg1) %>% exp() %>% round(digits = 5)
```

#### **Interpretation:**
+ The odds of having hypertension if you're a female is expected to increase by roughly 4%. 

+ For a one unit increase in age you would expect your odds of having hypertension to increase by roughly 7%. 

+ When you move from American Mexican race to Other Hispanic, the odds of having hypertension increases by roughly 25%.

+ When you move from American Mexican race to Non-Hispanic White, the odds of having hypertension increases by roughly 13%.

+ When you move from American Mexican race to Non-Hispanic Black, the odds of having hypertension is expected to be roughly 2.07 times higher.

+ When you move from American Mexican race to Other Race including multi-race, the odds of having hypertension increases by roughly 79%.

+ When you move from smoking to not smoking, the odds of having hypertension decreases by roughly 13%.

+ When you move from having diabetes to not having diabetes, the odds of having hypertension decreases by roughly 68%.

+ When you move from having diabetes to borderline having diabetes, the odds of having hypertension decreases by roughly 34%.

+ When you move from taking insulin to not taking insulin, the odds of having hypertension increases by roughly 38%.

+ When you move from having chronic kidney disease to not having chronic kidney disease, the odds of having hypertension decreases by roughly 78%.

+ For a one unit increase in BMI you would expect your odds of having hypertension to increase by roughly 8%. 

+ For a one unit increase in physical activity you would expect your odds of having hypertension to increase by roughly 9%. 

+ For a one unit increase in alcohol you would expect your odds of having hypertension to increase roughly by 7%.

+ When you move from never using salt to rarely, the odds of having hypertension decreases by roughly 25%.

+ When you move from never using salt to occasionally, the odds of having hypertension decreases by roughly 12%.

+ When you move from never using rarely to very often, the odds of having hypertension decreases by roughly 28%.

+ For a one unit increase in potassium you would expect your odds of having hypertension to be equally likely.


**Variables that decreases the odds of Hypertension **


+ Not Smoking 
+ Not Having Diabetes
+ Borderline Having Diabetes
+ Not Having Chronic Kidney Disease
+ Use of Salt in Food Prep


#### Key findings:

+ Salt regardless of frequency intake seems to decrease the odds of Hypertension
+ Variables that increase the odds of hypertension include: being female, a unit increase in age, race (especially non-hispanic black), not taking insulin, a unit increase in BMI and a unit increase in physical activity.
+ variables that decrease the odds of hypertension include: not smoking, not having diabetes, having borderline diabetes, not having chronic kidney disease, unit increase in alcohol (2015-2016 data only) and the use of salt in food prep.




**Plotting estimates of logistic regression models**

```{r}
#2015-2016 All Data
suppressMessages(library(sjPlot))
suppressMessages(library(sjlabelled))
suppressMessages(library(sjmisc))

theme_set(theme_sjplot())
plot_model(log_reg_all_dat, vline.color = "black", sort.est = TRUE)
plot_model(log_reg_all_dat, show.values = TRUE, value.offset = .3, sort.est = T)
```

```{r}
#2015-2016 Reduced Data
theme_set(theme_sjplot())
plot_model(log_reg, vline.color = "black", sort.est = TRUE)
plot_model(log_reg, show.values = TRUE, value.offset = .3, sort.est = T)
```


```{r}
#2009-2010 All Data
theme_set(theme_sjplot())
plot_model(log_reg_all, vline.color = "black", sort.est = TRUE)
plot_model(log_reg_all, show.values = TRUE, value.offset = .3, sort.est = T)
```

```{r}
#2009-2010 Reduced Data
theme_set(theme_sjplot())
plot_model(log_reg1, vline.color = "black", sort.est = TRUE)
plot_model(log_reg1, show.values = TRUE, value.offset = .3, sort.est = T)
```





**Residuals**


```{r}
# All data 2015-2016
plot(log_reg_all_dat)
```

```{r}
# 2015-1016 All Data
library(arm)
binnedplot(fitted(log_reg_all_dat), 
           residuals(log_reg_all_dat, type = "response"), 
           nclass = NULL, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")

# The grey lines represent  ±  2 SE bands, which we would expect to contain about 95% of the observations.
# https://bookdown.org/jefftemplewebb/IS-6489/logistic-regression.html 
```

```{r}
# 2015-2016 Reduced data
binnedplot(fitted(log_reg), 
           residuals(log_reg, type = "response"), 
           nclass = NULL, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")
```


```{r}
prob <- predict(log_reg, type="response")
pred <- factor(ifelse(prob < .5, "No", "Yes"))
table(hyp1$Hypertension, pred)
```




**Smoothing Spline For BMI and Age**

```{r}
# Hypertension vs BMI
library(splines)

f1 <- smooth.spline(hyp1$BMI,hyp1$Hypertension, cv = T)
# cross-validation with non-unique 'x' values seems doubtful

ggplot(data=hyp1, aes(x=BMI, y=Hypertension))+ geom_point() +
geom_line(aes(y=fitted(f1)), color="green")

f2 <- glm(Hypertension ~ bs(BMI,df=5), data=hyp1, family = binomial)
attr(bs(hyp1$BMI, df=5), "knots")

ggplot(data=hyp1, aes(x=BMI, y=Hypertension)) + geom_point() + 
  geom_line(aes(y=fitted(f2)), color= "blue") + geom_vline(xintercept = attr(bs(hyp1$BMI, df=5), "knots"), colour = "magenta")
```

```{r}
library(gam)
gam_logistic = gam(Hypertension ~ s(BMI,df = 3),
                   family = binomial, data = hyp1)
plot(gam_logistic, se = TRUE, col = "green")
```


**Random Forest**

```{r}
suppressMessages(library(randomForest))
rf <- randomForest(Hypertension ~ .,data=hyp1, importance = T, ntree = 1000)
#The response has five or fewer unique values.  Are you sure you want to do regression?
rf
varImpPlot(rf)


#Decision Trees

suppressMessages(library(tree))
tree <- tree(Hypertension ~ BMI + Age, data=hyp1)
summary(tree)
plot(tree)
text(tree, cex=.5, pretty=0)

library(rpart)
f1 <- rpart(Hypertension ~ ., data=hyp1)
plot(f1)
text(f1, cex=.5, pretty = 0)
```



